The problem and promise of scale in multilayer networks

### 2020-01-09 paper/manuscript.Rmd
Small edits to MS

## Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has become a central consideration in the way we understand landscape ecology and animal space use. Multilayer networks promise the integration of monolayer animal social networks with the complexity and importance of animal movement and space use in heterogenous landscapes. Despite the complex interplay between social networks and how animals use space, there remains an important biological and methodological gap in our understanding of an animals perception of spatial scale, e.g., grain and extent, can affect multilayer network dynamics. Here, we test how multilayer network metrics (e.g., versatility, multi-degree, similarity) are sensitive to variation in an animals perception of scale. We will simulate animal movement data coupled with emergent social networks on increasingly heterogenous landscapes across spatial scales. Furthermore, we review specific examples of spatially explicit data types to include in multilayer networks and explore the relevant nuances of these data. *Our simulations will highlight the need to carefully select scales for measuring spatial processes and appropriately weight layers and edges representing different types of data.* Effective integration of spatial processes, including biologically meaningful scales, within the context of animal social networks is an emerging area of research; our contribution will uniquely incorporate landscape ecology theory as a way to link different scales of social and spatial processes in a multilayer network. Based on social network and landscape ecology theory as well as our simulations we demonstrate potential for interconnectedness among individual animals and the landscapes they occupy.

## Introduction
Our understandings of ecological pattern and process are directly linked to the scale at which they are observed (Levin 1992; Allen et al 2015). Animals and landscapes are influenced across scales by higher level processes, such as climate and productivity (Field et al 2009) and lower level processes, such as species interactions and resource availability (Legendre 1993; Chave 2013). The effect of scale in ecology has been explored through species distributions (Elith et al 2009) habitat selection (Mayor et al 2007) and food webs (Sugihara et al 1989). Scale can be considered for all types of measured quantities9 including spatial and temporal variables, as well as behavioural variables including sociality.

Sociality exists across scales (Whitehead 2008). For example, grooming and mating relationships requires close proximity between conspecifics and occur over brief time periods (e.g. Carter et al. 2015), whereas social association represents shared space use by members of the same social group (Franks et al. 2010). Further, home range overlap among individuals is a coarse scale of sociality, where individuals share space over larger spatial and temporal scales (e.g. Piza-Roca et al. 2018). Shared space is a requirement for most types of social interactions, however, it remains unclear whether coarser scales of sociality, including social association and home range overlap, vary predictably with social interaction (Castles et al. 2014; Farine 2015). Social scale can be defined as the type of relationship observed (Farine 2015) and these relationships can be explicitly defined and measured in social network analyses (Farine et al. 2015).

Animal social network analysis are a well-developed tool used to measure the relationships of individuals and organization of social systems (Krause et al 2009; Wey et al. 2008; Croft et al. 2008). Social network analyses provide insight into interspecific interactions and the structure of social communities and social network properties can influences population dynamics and evolutionary processes (Pinter-Wollman et al. 2014; Kurvers et al. 2014). Nevertheless, despite the power of traditional social network analysis, it typically considers a single scale of sociality in a given network, drastically simplifying the complexity of animal social systems (Finn et al. 2019).

Unlike traditional social networks, multilayer networks explicitly consider social systems across contexts, including scale-dependent contexts (Pilosof et al. 2017). Multilayer networks are made up of multiple layers, each representing different classes of individuals (e.g. male or female), types of behaviours (e.g. grooming or mating), spatial areas (e.g. local or regional), or temporal windows (e.g. daily or seasonal) (Kivela et al 2014; Porter 2018). Multilayer networks are relatively novel to describing animal behaviour studies (Finn et al. 2019; Silk et al. 2018), although they have been used to describe multidimensional human social systems, complex transportation networks, and the complete neural network of an organism (Silk et al. 2018). The role of scale within a multilayer social network context remains unexplored, but provides a unique opportunity to develop novel understanding of variation in the temporal or spatial scales associated with social systems.
Here, we discuss and integrate variation in temporal and spatial scales within a multilayer network framework. We borrow from behavioural, landscape, and spatial ecology to build a conceptual and analytical framework for scale-dependent multilayer networks and we apply this framework seasonally variable social system case study of caribou (Rangifer tarandus) in Newfoundland, Canada (Peignier et al. 2019). We use this case study to address two principal questions associated with scale-dependent multilayer networks. First, to what degree does the spatial scale of landscapes, temporal windows of analysis, and social scale of relationships, influence our ability to interpret complex social systems? Second, can variation in spatial, temporal, and social scale improve predictions associated with seasonal differences in resource availability and social association. We conclude with perspectives for further development and application of scale in multilayer networks by discussing the problem and promise of scale from the perspective of social systems (Castles et al. 2014; Farine 2015) and landscape ecology (Schneider 2009).

@@

### Caribou socioecology
Here, we use caribou (Rangifer tarandus) as a case study to develop scale-dependent multilayer networks based on temporal (seasonal), spatial (habitat type), and social (visual and auditory perception). Caribou are gregarious ungulates with fission-fusion dynamics (Lesmerises et al. 2018) that display temporal (Peignier et al. 2019) and spatial (Webber et al. unpublished data) variation in social networks. In winter, caribou dig holes in the snow, termed craters to access forage (Bergerud 1974). As a result, caribou tend to occupy and re-use craters once they are established. The average area of craters dug by caribou in Newfoundland was 0.41 m2 (SD = 0.48; Mayor et al. 2009) and crater density, which varies based on snow condition, depth, and local caribou density, can range from 366–1980 craters/ha (Bergerud 1974). There is therefore considerably less access to forage than when the landscape is snow-free. Importantly, craters exist on the landscape at multiple scales: the carter scale (a single crater), the feeding area scale (multiple craters in close proximity), and the winter range scale (all craters within an individual’s range) (Mayor et al. 2009). The distribution of craters on the landscape is therefore heterogeneous and access to vegetation in winter is highly variable among individual caribou.

### Caribou location data
We used GPS location data collected from Fogo Island caribou in winter 2018 (1 January – 16 March). Adult female caribou were immobilized and fitted with global positioning system (GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars, 1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas et al. (2010). We did not collar all female caribou in the herd, however, we assumed that our sample of collared animals was random. Although associations between collared and uncollared animals were unrecorded, we assumed that our networks (see below) were unbiased representations of the relative degree of social association among all caribou.

>>>>

## Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has become a central consideration in the way we understand landscape ecology and animal space use. Multilayer networks promise the integration of monolayer animal social networks with the complexity and importance of animal movement and space use in heterogeneous landscapes. Despite the complex interplay between social networks and how animals use space, there remains an important biological and methodological gap in our understanding of an animal’s perception of spatial scale, e.g., grain and extent, can affect multilayer network dynamics. Here, we test how multilayer network metrics (e.g., versatility, multi-degree, similarity) are sensitive to variation in an animal’s perception of scale. We will simulate animal movement data coupled with emergent social networks on increasingly heterogenous landscapes across spatial scales. Furthermore, we review specific examples of spatially explicit data types to include in multilayer networks and explore the relevant nuances of these data. *Our simulations will highlight the need to carefully select scales for measuring spatial processes and appropriately weight layers and edges representing different types of data.* Effective integration of spatial processes, including biologically meaningful scales, within the context of animal social networks is an emerging area of research; our contribution will uniquely incorporate landscape ecology theory as a way to link different scales of social and spatial processes in a multilayer network. Based on social network and landscape ecology theory as well as our simulations we demonstrate potential for interconnectedness among individual animals and the landscapes they occupy.

## Introduction
Our understandings of ecological pattern and process are directly linked to the scale at which they are observed (Levin 1992; Allen et al 2015). Animals and landscapes are influenced across scales by higher level processes, such as climate and productivity (Field et al 2009) and lower level processes, such as species interactions and resource availability (Legendre 1993; Chave 2013). The effect of scale in ecology has been explored through species distributions (Elith et al 2009) habitat selection (Mayor et al 2007) and food webs (Sugihara et al 1989). Scale can be considered for all types of measured quantities including spatial and temporal variables, as well as behavioural variables, including sociality.

Sociality exists across scales (Whitehead 2008). For example, grooming and mating relationships require close proximity between conspecifics and occur over brief time periods (e.g. Carter et al. 2015), whereas social association represents shared space use by members of the same social group (Franks et al. 2010). Further, home range overlap among individuals is a coarse scale of sociality, where individuals share space over larger spatial and temporal scales (e.g. Piza-Roca et al. 2018). Shared space is a requirement for most types of social interactions. However, it remains unclear whether coarser scales of sociality, including social association and home range overlap, vary predictably with social interaction (Castles et al. 2014; Farine 2015). Social scale can be defined as the type of relationship observed (Farine 2015) and these relationships can be explicitly defined and measured in social network analyses (Farine et al. 2015).

Animal social network analysis is a well-developed tool used to measure the relationships of individuals and organization of social systems (Krause et al 2009; Wey et al. 2008; Croft et al. 2008). Social network analyses provide insight into interspecific interactions and the structure of social communities and social network properties can influences population dynamics and evolutionary processes (Pinter-Wollman et al. 2014; Kurvers et al. 2014). Nevertheless, despite the power of traditional social network analysis, it typically considers a single scale of sociality in a given network, drastically simplifying the complexity of animal social systems (Finn et al. 2019).

Unlike traditional social networks, multilayer networks explicitly consider social systems across contexts, including scale-dependent contexts (Pilosof et al. 2017). Multilayer networks are made up of multiple layers, each representing different classes of individuals (e.g. male or female), types of behaviours (e.g. grooming or mating), spatial areas (e.g. local or regional), or temporal windows (e.g. daily or seasonal) (Kivela et al 2014; Porter 2018). Multilayer networks are relatively novel to describing animal behaviour studies (Finn et al. 2019; Silk et al. 2018), although they have been used to describe multidimensional human social systems, complex transportation networks, and the complete neural network of an organism (Silk et al. 2018). The role of scale within a multilayer social network context remains unexplored but provides a unique opportunity to develop novel understanding of variation in the temporal or spatial scales associated with social systems.

Here, we discuss and integrate variation in temporal and spatial scales within a multilayer network framework. We borrow from behavioural, landscape, and spatial ecology to build a conceptual and analytical framework for scale-dependent multilayer networks and we apply this framework seasonally variable social system case study of caribou (*Rangifer tarandus*) in Newfoundland, Canada (Peignier et al. 2019). We use this case study to address two principal questions associated with scale-dependent multilayer networks. First, to what degree does the spatial scale of landscapes, temporal windows of analysis, and social scale of relationships, influence our ability to interpret complex social systems? Second, can variation in spatial, temporal, and social scale improve predictions associated with seasonal differences in resource availability and social association? We conclude with perspectives for further development and application of scale in multilayer networks by discussing the problem and promise of scale from the perspective of social systems (Castles et al. 2014; Farine 2015) and landscape ecology (Schneider 2009).

@@

### Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer networks based on temporal (seasonal), spatial (habitat type), and social (visual and auditory perception). Caribou are gregarious ungulates with fission-fusion dynamics (Lesmerises et al. 2018) that display temporal (Peignier et al. 2019) and spatial (Webber et al. unpublished data) variation in social networks. In winter, caribou dig holes in the snow, termed craters to access forage (Bergerud 1974). As a result, caribou tend to occupy and re-use craters once they are established. The average area of craters dug by caribou in Newfoundland was 0.41 m2 (SD = 0.48; Mayor et al. 2009) and crater density, which varies based on snow condition, depth, and local caribou density, can range from 366–1980 craters/ha (Bergerud 1974). There is therefore considerably less access to forage than when the landscape is snow-free. Importantly, craters exist on the landscape at multiple scales: the crater scale (a single crater), the feeding area scale (multiple craters in close proximity), and the winter range scale (all craters within an individual’s range) (Mayor et al. 2009). The distribution of craters on the landscape is therefore heterogeneous and access to vegetation in winter is highly variable among individual caribou.

### Caribou location data
We used GPS location data collected from Fogo Island caribou in winter 2018 (1 January – 16 March). Adult female caribou were immobilized and fitted with global positioning system (GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars, 1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas et al. (2010). We did not collar all female caribou in the herd. However, we assumed that our sample of collared animals was random. Although associations between collared and uncollared animals were unrecorded, we assumed that our networks (see below) were unbiased representations of the relative degree of social association among all caribou.
